Data Descriptor: Corresponding author journal publishing activities 2014-2018

### 2019-10-19 analysis/paper.Rmd
update data descriptor

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(scatterpie)
library(knitr)
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)

@@

1. Initial exploration about how corresponding authorships are represented in the WoS-KB ([`001_kb_rp_pretest.Rmd`](analysis/001_kb_rp_pretest.Rmd)) 
2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)). 
3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).

@@

- `journal_publisher_14_18`: Global journal and publisher data 2014-18
- `rp_jn_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18

@@

```{r}
rp_df <- readr::read_csv("../data/rp_jn_14_18.csv")
```

The file named `rp_jn_14_18` provides a breakdown of journal information aggregated by  countries of affiliation. Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration. The following table  describes the variables in detail. Overall, 99 % of records representing original articles and reviews in the period 2014 - 2018 had affiliation information for corresponding authors at the country-level.

@@

| Variable           | Description                                                                                                                                                                      | Source                                                                                                                                                     |
| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | :--------------------------------------------------------------------------------------------------------------------------------------------------------- |
| `issn_wos`         | ISSN, a standardized journal id.                                                                                                                                                 | KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                |
| `country_code`     | Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           | KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                              |
| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                                                             | KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                              |
| `articles`         | Number of original articles and reviews published. Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally. | KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`, `wos_b_2019.d_items_authors_institutions.inst_countrycode` and `wos_b_2019.items.pubyear` |
| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       | Crossref                                                                                                                                                   |
| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      | Crossref                                                                                                                                                   |
| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?                                                                                             | ISSN-GOLD_OA List 3.0                                                                                                                              |
| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             | CIEPS                                                                                                                                                      |

@@

Using `rp_14_18` the publisher market shares can be further broken down by country affiliation from corresponding authors. The following figure presents a global map of scholarly journal publishing represented in the Web of Science. Publications from institutions based in the European Union were summarized as "EU28". Together, all EU 28 member states constituted the largest market for journal publishing between 2014-18.

```{r}
my_df <- readr::read_csv("../data/rp_jn_14_18.csv")
#'  transform to eu
eu <-
  c(
    "AUT",
    "BEL",
    "BGR",
    "CYP",
    "CZE",
    "DEU",
    "DNK",
    "ESP",
    "EST",
    "FIN",
    "FRA",
    "GBR",
    "GRC",
    "HRV",
    "HUN",
    "IRL",
    "ITA",
    "LTU",
    "LUX",
    "LVA",
    "MLT",
    "NLD",
    "POL",
    "PRT",
    "ROU",
    "SVK",
    "SVN",
    "SWE"
  )
my_df %>%
  mutate(eu = case_when(country_code %in% eu ~ "EU28")) %>%
  mutate(countries = ifelse(eu %in% "EU28", "EU28", country_code)) -> df_to_plot
#' top 5 publisher closed
my_df %>%
 # filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(articles)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher -> top_6
#' data transformation to prepare plotting of corresponding authors' country of affilation
df_to_plot %>%
  mutate(fct_publisher = ifelse(publisher %in% top_6, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  group_by(fct_publisher, countries) %>%
  summarize(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(fct_publisher = factor(fct_publisher, levels = c(top_6[!is.na(top_6)], "Others"))) %>%
  mutate(countries = fct_reorder(countries, n)) -> plot_df
#' add country lat lon
#' eu28 frankfurt main is the center
eu_df <-
  data_frame(
    latitude = 50.6393264,
    longitude = 8.40556572,
    name = "EU28",
    iso3c = "EU28"
  )
library(countrycode)
gc <- readr::read_tsv("../data/geocodes.csv") %>%
  # add iso code used in kb database
  mutate(iso3c = countrycode::countrycode(country, "iso2c", "iso3c")) %>%
  bind_rows(eu_df) %>%
  filter(!iso3c %in% eu)

@@

The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations. The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher. The figure indicates, which countries were of particularly importance to a publisher. Furthermore, it suggests large variations across publishers. Take for instance the publisher Royal Society of Chemistry. Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.

```{r}
top_10_publisher <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(articles)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher
top_10_countries <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(country_code) %>%
  summarise(n = sum(articles)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  .$country_code
plot_df <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  mutate(fct_publisher = ifelse(publisher %in% top_10_publisher, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  mutate(fct_countrcode = ifelse(country_code %in% top_10_countries, country_code, "Others")) %>%
  mutate(fct_countrcode = factor(fct_countrcode, levels = c(top_10_countries, "Others"))) %>%
  group_by(fct_publisher, fct_countrcode) %>%
  summarize(n = sum(articles)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()
#' plot
ggplot(plot_df, aes(forcats::fct_rev(fct_countrcode), prop, fill = prop)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  facet_wrap(~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
  scale_fill_viridis_c(
    "Country share per publisher (in %)",
    labels = scales::percent_format(accuracy = 5L),
    option = "C"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L))  +
  labs(
    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
    x = "Country of Affiliation (Top 10)",
    y = "Share of Original Articles and Reviews per Publisher",
    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
    subtitle = "Corresponding authors' countries of affiliation per publisher."
  ) +
  theme_minimal(base_family = "Roboto", base_size = 16) +
  theme(plot.margin = margin(30, 10, 30, 10)) +
  theme(panel.grid.minor.y = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size =  9)
  ) -> p
#' save as png
ggsave(
  "figure/publisher_share.png",
  p,
  dpi = 450,
  height = 13,
  width = 17
)
```

@@

>>>>

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(scatterpie)
library(knitr)
library(scales)
library(countrycode)
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)

@@

1. Initial exploration about how corresponding authorships are represented in the WoS-KB ([`001_kb_rp_pretest.Rmd`](analysis/001_kb_rp_pretest.Rmd)) 
2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)). 
3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018 ([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).

@@

- `journal_publisher_14_18`: Global journal and publisher data 2014-18
- `rp_publisher_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18

@@

```{r}
rp_df <- readr::read_csv("../data/rp_publisher_14_18.csv")
```

The dataset named `rp_publisher_14_18` provides a breakdown of journal information aggregated by  countries of affiliation. Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration. The following table  describes the variables in detail. Overall, 99 % of records representing original articles and reviews in the period 2014 - 2018 had affiliation information for corresponding authors at the country-level.

@@

|Variable           |Description                                                                                                                                 |Source                                                                        |
|:------------------|:-------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------|
|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                      |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode` |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                        |KB Web of Science: `wos_b_2019.items.pubyear`                                 |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref |Crossref                                                                      |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                        |ISSN GOLD OA List V3                                                          |
|`n_publications`   |Number of original articles and reviews published                                                                                           |Aggregate data                                                                |
|`n_journals`       |Number of distinct journals with corresponding publication output                                                                           |Aggregate data

@@

Using `rp_publisher_14_18.csv` the publisher market shares can be further broken down by country affiliation from corresponding authors. The following figure presents a global map of scholarly journal publishing represented in the Web of Science. Publications from institutions based in the European Union were summarized as "EU28". Together, all EU 28 member states constituted the largest market for journal publishing between 2014-18.

```{r}
#'  transform to eu
eu <-
  c(
    "AUT",
    "BEL",
    "BGR",
    "CYP",
    "CZE",
    "DEU",
    "DNK",
    "ESP",
    "EST",
    "FIN",
    "FRA",
    "GBR",
    "GRC",
    "HRV",
    "HUN",
    "IRL",
    "ITA",
    "LTU",
    "LUX",
    "LVA",
    "MLT",
    "NLD",
    "POL",
    "PRT",
    "ROU",
    "SVK",
    "SVN",
    "SWE"
  )
rp_df %>%
  mutate(eu = case_when(country_code %in% eu ~ "EU28")) %>%
  mutate(countries = ifelse(eu %in% "EU28", "EU28", country_code)) -> df_to_plot
#' top 5 publisher closed
rp_df %>%
 # filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(n_publications)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher -> top_6
#' data transformation to prepare plotting of corresponding authors' country of affilation
df_to_plot %>%
  mutate(fct_publisher = ifelse(publisher %in% top_6, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  group_by(fct_publisher, countries) %>%
  summarize(n = sum(n_publications)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(fct_publisher = factor(fct_publisher, levels = c(top_6[!is.na(top_6)], "Others"))) %>%
  mutate(countries = fct_reorder(countries, n)) -> plot_df
#' add country lat lon
#' eu28 frankfurt main is the center
eu_df <-
  tibble(
    latitude = 50.6393264,
    longitude = 8.40556572,
    name = "EU28",
    iso3c = "EU28"
  )
library(countrycode)
gc <- readr::read_tsv("../data/geocodes.csv") %>%
  # add iso code used in kb database
  mutate(iso3c = countrycode::countrycode(country, "iso2c", "iso3c")) %>%
  bind_rows(eu_df) %>%
  filter(!iso3c %in% eu)

@@

The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations. The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher. The figure indicates, which countries were of particularly importance to a publisher. Furthermore, it suggests large variations across publishers. Take for instance the publisher Royal Society of Chemistry. Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.

```{r}
top_10_publisher <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(publisher) %>%
  summarise(n = sum(n_publications)) %>%
  arrange(desc(n)) %>%
  head(11) %>%
  .$publisher
top_10_countries <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  group_by(country_code) %>%
  summarise(n = sum(n_publications)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  .$country_code
plot_df <- rp_df %>%
  filter(oa_journal == FALSE) %>%
  mutate(fct_publisher = ifelse(publisher %in% top_10_publisher, publisher, "Others")) %>%
  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
  mutate(fct_countrcode = ifelse(country_code %in% top_10_countries, country_code, "Others")) %>%
  mutate(fct_countrcode = factor(fct_countrcode, levels = c(top_10_countries, "Others"))) %>%
  group_by(fct_publisher, fct_countrcode) %>%
  summarize(n = sum(n_publications)) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()
#' plot
ggplot(plot_df, aes(forcats::fct_rev(fct_countrcode), prop, fill = prop)) +
  coord_flip() +
  geom_bar(stat = "identity") +
  facet_wrap(~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
  scale_fill_viridis_c(
    "Country share per publisher (in %)",
    labels = scales::percent_format(accuracy = 5L),
    option = "C"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L))  +
  labs(
    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
    x = "Country of Affiliation (Top 10)",
    y = "Share of Original Articles and Reviews per Publisher",
    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
    subtitle = "Corresponding authors' countries of affiliation per publisher."
  ) +
  theme_minimal(base_family = "Roboto", base_size = 16) +
  theme(plot.margin = margin(30, 10, 30, 10)) +
  theme(panel.grid.minor.y = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size =  9)
  ) -> p
#' save as png
ggsave(
  "figure/publisher_share.png",
  p,
  dpi = 450,
  height = 13,
  width = 17
)
```

@@

This work uses Web of Science data by Clarivate Analytics provided by the German Competence Center for Bibliometrics for research purposes.

This work is supported by the Federal Ministry of Education and Research of Germany (BMBF) in the framework Quantitative Research on the Science Sector (Project: "OAUNI Entwicklung und Einflussfaktoren des Open-Access-Publizierens an Universitäten in Deutschland", Förderkennzeichen: 01PU17023A).
